---
title: "VAR Predictability Benchmark"
subtitle: "Meta Regression"
author: 
 - name: Björn S. Siepe
   orcid: 0000-0002-9558-4648
   affiliations: University of Marburg
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    number-sections: true
    theme: cosmo
    code-fold: true
    code-tools: true
    code-summary: "Show the code"
    fig-width: 7
    fig-height: 4.5
    fig-align: "center"
    embed-resources: true
execute:
  message: false
  warning: false
  eval: true
params:
  refit_models: false
---


# Preparation

We first load all relevant packages and data.

```{r}
#| label: setup
library(here)
library(tibble)
library(dplyr)
library(brms)
library(pander)
library(parameters)
library(bayesplot)

seed <- 395013

# load data
df_meta <- readRDS(here::here("output", "meta", "combined.rds"))

# Make passes TRUE when the upstream combined results changed
# and direct renders reuse cached fits
# params only exists during knitr rendering, fall back to FALSE interactively
refit_models <- isTRUE(if (exists("params")) params$refit_models else FALSE)

```


We then need to create unique identifiers for each individual, ensuring that the same id number is not mistaken as the same person across different datasets. We do this by combining the dataset name and the person id.

```{r}
#| label: create-unique-id
df_meta <- df_meta |>
  dplyr::mutate(unique_id = paste0(dataset_id, "_", id))
```

```{r}
#| label: data-relative
# the mean model is the reference (rel_mse = 1 by construction), so it is not a row here
df_rel <- df_meta |>
  dplyr::filter(set == "oos", model != "mean", is.finite(rel_mse), rel_mse > 0) |>
  dplyr::mutate(model = droplevels(model))
```

```{r}
#| label: excluded-relative
# rows without a defined relative MSE, by reason (reason_na from relative_mse() in 03)
df_meta |>
  dplyr::filter(set == "oos", model != "mean", !is.finite(rel_mse)) |>
  dplyr::count(model, reason_na) |>
  pander::pander(caption = "OOS person rows excluded from the relative MSE models.")
```

# Meta Regression for Relative MSE


## Log-Normal Models


### Fixed-Effect for Model, Random-Effect for Dataset (M1)
We now fit a simple meta regression model for the relative MSE with a log-normal distribution and default priors. 

```{r}
#| label: meta-relmse-m1
model_path <- here::here("output", "analysis", "fit_relmse_m1.rds")

priors_rel <- c(
  prior(normal(0, 1), class = b),
  prior(exponential(2), class = sd),
  prior(normal(-1, 1), class = b, dpar = sigma)
)

if (refit_models || !file.exists(model_path)) {
  fit_relmse_m1 <- brm(
    # no intercept: the mean model is the reference and already divided out of rel_mse,
    # so each coefficient is that model's log relative MSE
    bf(rel_mse ~ 0 + model + (1 | dataset_id) + (1 | dataset_id:model) + (1 | unique_id),
       sigma ~ 0 + model),
    data = df_rel,
    family = lognormal(),
    prior = priors_rel,
    chains = 4,
    cores = 4,
    iter = 3000,
    warmup = 1000,
    seed = seed
  )

  saveRDS(fit_relmse_m1, model_path)
} else {
  fit_relmse_m1 <- readRDS(model_path)
}
```

These priors are: 

```{r}
#| label: priors-rmse-lognormal
brms::prior_summary(fit_relmse_m1) |>
  dplyr::select(prior, class, coef, group) |>
  pander::pander(caption = "Priors for the log-normal meta regression model for RMSE.")
```

```{r}
#| label: prior-predictive-m1
model_path <- here::here("output", "analysis", "fit_relmse_m1_prior.rds")

if (refit_models || !file.exists(model_path)) {
  fit_relmse_m1_prior <- update(fit_relmse_m1, sample_prior = "only", seed = seed)
  saveRDS(fit_relmse_m1_prior, model_path)
} else {
  fit_relmse_m1_prior <- readRDS(model_path)
}

bayesplot::pp_check(fit_relmse_m1_prior, type = "dens_overlay", ndraws = 50) +
  ggplot2::scale_x_log10()
```

Here's the summary of the model: 
```{r}
#| label: summary-rmse-lognormal
parameters::model_parameters(fit_relmse_m1, test = "pd")
```

Here's a traceplot of some main parameters: 

```{r}
#| label: traceplot-rmse-lognormal
bayesplot::mcmc_trace(fit_relmse_m1, regex_pars = "^b_")
```

Here's a posterior predictive check of the model:
```{r}
#| label: ppcheck-rmse-lognormal
bayesplot::pp_check(fit_relmse_m1, type = "dens_overlay_grouped", group = "model", ndraws = 50) +
  ggplot2::scale_x_log10()
```


### M1 + Training Length by Model (M2)

```{r}
#| label: moderator-training-length
source(here::here("scripts", "engine", "relative.R"))
features <- readRDS(here::here("output", "analysis", "features.rds"))

# log training length split into the dataset mean (between) and the person's deviation from it (within)
train_length <- features$person |>
  dplyr::mutate(log_n = log(n_train_first)) |>
  split_between_within(log_n, dataset_id) |>
  dplyr::select(dataset_id, id, log_n_between, log_n_within)

df_rel <- dplyr::left_join(df_rel, train_length, by = c("dataset_id", "id"))
stopifnot(!anyNA(df_rel$log_n_within))
```

```{r}
#| label: meta-relmse-m2
model_path <- here::here("output", "analysis", "fit_relmse_m2.rds")

if (refit_models || !file.exists(model_path)) {
  fit_relmse_m2 <- brm(
    # one between and one within slope per model
    bf(rel_mse ~ 0 + model + model:log_n_between + model:log_n_within +
         (1 | dataset_id) + (1 | dataset_id:model) + (1 | unique_id),
       sigma ~ 0 + model),
    data = df_rel,
    family = lognormal(),
    prior = priors_rel,
    chains = 4,
    cores = 4,
    iter = 3000,
    warmup = 1000,
    seed = seed
  )

  saveRDS(fit_relmse_m2, model_path)
} else {
  fit_relmse_m2 <- readRDS(model_path)
}
```

```{r}
#| label: summary-relmse-m2
parameters::model_parameters(fit_relmse_m2, test = "pd")
```


### M2 + Autocorrelation by Model (M3)

```{r}
#| label: moderator-autocorrelation
# mean lag-1 autocorrelation of the items in the training window, split like training length
autocorrelation <- features$person |>
  dplyr::mutate(rho1 = rho1_train_mean) |>
  split_between_within(rho1, dataset_id) |>
  dplyr::select(dataset_id, id, rho1_between, rho1_within)

# rho1 is undefined for persons without observed lag pairs in the training window
df_m3 <- df_rel |>
  dplyr::left_join(autocorrelation, by = c("dataset_id", "id")) |>
  dplyr::filter(!is.na(rho1_within))
```

```{r}
#| label: meta-relmse-m3
model_path <- here::here("output", "analysis", "fit_relmse_m3.rds")

if (refit_models || !file.exists(model_path)) {
  fit_relmse_m3 <- brm(
    bf(rel_mse ~ 0 + model + model:log_n_between + model:log_n_within +
         model:rho1_between + model:rho1_within +
         (1 | dataset_id) + (1 | dataset_id:model) + (1 | unique_id),
       sigma ~ 0 + model),
    data = df_m3,
    family = lognormal(),
    prior = priors_rel,
    chains = 4,
    cores = 4,
    iter = 3000,
    warmup = 1000,
    seed = seed
  )

  saveRDS(fit_relmse_m3, model_path)
} else {
  fit_relmse_m3 <- readRDS(model_path)
}
```

```{r}
#| label: summary-relmse-m3
parameters::model_parameters(fit_relmse_m3, test = "pd")
```

# Sensitivity Analyses

## Student-t Likelihood (S1)

```{r}
#| label: meta-relmse-s1
model_path <- here::here("output", "analysis", "fit_relmse_s1.rds")

if (refit_models || !file.exists(model_path)) {
  fit_relmse_s1 <- brm(
    # M1 on the log scale with heavier tails, which downweights persons where a model fails badly
    bf(log(rel_mse) ~ 0 + model + (1 | dataset_id) + (1 | dataset_id:model) + (1 | unique_id),
       sigma ~ 0 + model),
    data = df_rel,
    family = student(),
    prior = priors_rel,
    chains = 4,
    cores = 4,
    iter = 3000,
    warmup = 1000,
    seed = seed
  )

  saveRDS(fit_relmse_s1, model_path)
} else {
  fit_relmse_s1 <- readRDS(model_path)
}
```

```{r}
#| label: prior-predictive-s1
model_path <- here::here("output", "analysis", "fit_relmse_s1_prior.rds")

if (refit_models || !file.exists(model_path)) {
  fit_relmse_s1_prior <- update(fit_relmse_s1, sample_prior = "only", seed = seed)
  saveRDS(fit_relmse_s1_prior, model_path)
} else {
  fit_relmse_s1_prior <- readRDS(model_path)
}

# the response is already log(rel_mse), so no log axis
bayesplot::pp_check(fit_relmse_s1_prior, type = "dens_overlay", ndraws = 50)
```

```{r}
#| label: summary-relmse-s1
parameters::model_parameters(fit_relmse_s1, test = "pd")
```

## Raw-Scale MSE (S4)

```{r}
#| label: meta-relmse-s4
model_path <- here::here("output", "analysis", "fit_relmse_s4.rds")

# all models including the mean model, so the reference is modelled instead of divided out
df_raw <- df_meta |>
  dplyr::filter(set == "oos", is.finite(gm_mse), gm_mse > 0)

# raw MSE on the [0, 1] scale is around exp(-3), not 1
priors_raw <- c(
  prior(normal(-3, 2), class = b),
  prior(exponential(2), class = sd),
  prior(normal(-1, 1), class = b, dpar = sigma)
)

if (refit_models || !file.exists(model_path)) {
  fit_relmse_s4 <- brm(
    bf(gm_mse ~ 0 + model + (1 | dataset_id) + (1 | dataset_id:model) + (1 | unique_id),
       sigma ~ 0 + model),
    data = df_raw,
    family = lognormal(),
    prior = priors_raw,
    chains = 4,
    cores = 4,
    iter = 3000,
    warmup = 1000,
    seed = seed
  )

  saveRDS(fit_relmse_s4, model_path)
} else {
  fit_relmse_s4 <- readRDS(model_path)
}
```

```{r}
#| label: prior-predictive-s4
model_path <- here::here("output", "analysis", "fit_relmse_s4_prior.rds")

if (refit_models || !file.exists(model_path)) {
  fit_relmse_s4_prior <- update(fit_relmse_s4, sample_prior = "only", seed = seed)
  saveRDS(fit_relmse_s4_prior, model_path)
} else {
  fit_relmse_s4_prior <- readRDS(model_path)
}

bayesplot::pp_check(fit_relmse_s4_prior, type = "dens_overlay", ndraws = 50) +
  ggplot2::scale_x_log10()
```

```{r}
#| label: summary-relmse-s4
parameters::model_parameters(fit_relmse_s4, test = "pd")
```

# Convergence Diagnostics

```{r}
#| label: convergence
# population-level and variance parameters only. the person-level effects are too many to list
convergence <- function(fit) {
  draws <- brms::as_draws_df(fit, variable = "^(b_|sd_|nu$)", regex = TRUE)
  summary <- posterior::summarise_draws(draws, "rhat", "ess_bulk", "ess_tail")
  nuts <- brms::nuts_params(fit)
  tibble::tibble(
    max_rhat = max(summary$rhat),
    min_ess_bulk = min(summary$ess_bulk),
    min_ess_tail = min(summary$ess_tail),
    divergences = sum(nuts$Value[nuts$Parameter == "divergent__"])
  )
}

list(M1 = fit_relmse_m1, M2 = fit_relmse_m2, M3 = fit_relmse_m3,
     S1 = fit_relmse_s1, S4 = fit_relmse_s4) |>
  lapply(convergence) |>
  dplyr::bind_rows(.id = "model") |>
  pander::pander(digits = 3, caption = "Convergence diagnostics per fitted model.")
```

# Session Info

```{r}
#| label: session-info
pander::pander(sessionInfo())
```
