---
title: "VAR Predictability Benchmark"
subtitle: "Meta Regression"
author: 
 - name: Björn S. Siepe
   orcid: 0000-0002-9558-4648
   affiliations: University of Marburg
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    number-sections: true
    theme: cosmo
    code-fold: true
    code-tools: true
    code-summary: "Show the code"
    fig-width: 7
    fig-height: 4.5
    fig-align: "center"
    embed-resources: true
execute:
  message: false
  warning: false
  eval: true
---


# Preparation

We first load all relevant packages and data.

```{r}
#| label: setup
library(here)
library(tibble)
library(dplyr)
library(brms)
library(pander)
library(parameters)
library(bayesplot)

seed <- 395013

# load data
df_meta <- readRDS(here::here("output", "meta", "combined.rds"))

# should all models be refit?
refit_models <- TRUE

```


We then need to create unique identifiers for each individual, ensuring that the same id number is not mistaken as the same person across different datasets. We do this by combining the dataset name and the person id.

```{r}
#| label: create-unique-id
df_meta <- df_meta |>
  dplyr::mutate(unique_id = paste0(dataset_id, "_", id))
```

# Meta Regression for RMSE

Before we conduct the meta-regression, we need to average the RMSE values across all variables for each person in a dataset. This is necessary because the RMSE values are calculated per variable, and we want to analyze the overall performance of each model for each individual. We also select only the out-of-sample RMSE values, as these are the most relevant for evaluating model performance.

```{r}
#| label: average-rmse
df_meta <- df_meta |>
  # summarise across all variables to keep all moderators in the dataset
  dplyr::group_by(dplyr::across(-c(rmse, R2))) |>
  dplyr::summarise(mean_rmse = mean(rmse, na.rm = TRUE), .groups = "drop")

# select oos values
df_meta <- df_meta |>
  dplyr::filter(set == "oos")
```

We then need to check the amount of true zeroes, because the log-normal distribution cannot handle zeroes. If there are too many zeroes, we may need to consider a different distribution or transformation.

```{r}
#| label: check-zeroes
zeroes <- sum(df_meta$mean_rmse == 0, na.rm = TRUE)
```

There are `r zeroes` true zeroes in the RMSE values. We can thus proceed with the log-normal model, and exclude this individual. 

```{r}
#| label: remove-zero

df_meta <- df_meta |> 
  filter(mean_rmse != 0)
```


## Log-Normal Models


### Fixed-Effect for Model, Random-Effect for Dataset (M1)
We now fit a simple meta regression model for RMSE with a log-normal distribution and default priors. 

```{r}
#| label: meta-rmse-lognormal
model_path <- here::here("output", "analysis", "fit_rmse_lognormal.rds")

if (refit_models || !file.exists(model_path)) {
  fit_rmse_lognormal <- brm(
    mean_rmse ~ 1 + model + (1 | dataset_id) + (1 | unique_id),
    data = df_meta,
    family = lognormal(),
    chains = 4,
    cores = 4,
    iter = 3000,
    warmup = 1000,
    seed = seed
  )

  saveRDS(fit_rmse_lognormal, model_path)
} else {
  fit_rmse_lognormal <- readRDS(model_path)
}
```

These priors are: 

```{r}
#| label: priors-rmse-lognormal
brms::get_prior(fit_rmse_lognormal)|>
  dplyr::select(prior, class, coef, group) |>
  pander::pander(caption = "Priors for the log-normal meta regression model for RMSE.")
```

Here's the summary of the model: 
```{r}
#| label: summary-rmse-lognormal
parameters::model_parameters(fit_rmse_lognormal, test = "pd")
```

Here's a traceplot of some main parameters: 

```{r}
#| label: traceplot-rmse-lognormal
bayesplot::mcmc_trace(fit_rmse_lognormal, pars = c("b_Intercept", "b_modellocf", "b_modeltrend", "b_modelri", "b_modelar", "b_modelvar", "b_modelml_ar", "b_modelml_var"))
```

Here's a posterior predictive check of the model:
```{r}
#| label: ppcheck-rmse-lognormal
bayesplot::pp_check(fit_rmse_lognormal, type = "dens_overlay", ndraws = 100)
```


### M1 + Moderator for Number of Timepoints (M2)

```{r}
#| label: meta-rmse-lognormal-timepoint

model_path <- here::here("output", "analysis", "fit_rmse_lognormal_timepoint.rds")

if (refit_models || !file.exists(model_path)) {
  fit_rmse_lognormal_timepoint <- brm(
    mean_rmse ~ 1 + model + n_valid_person + (1 | dataset_id) + (1 | unique_id),
    data = df_meta,
    family = lognormal(),
    chains = 4,
    cores = 4,
    iter = 3000,
    warmup = 1000,
    seed = seed
  )

  saveRDS(fit_rmse_lognormal_timepoint, model_path)
} else {
  fit_rmse_lognormal_timepoint <- readRDS(model_path)
}
```


### M2 + Interaction of Model with Number of Timepoints (M2)

```{r}
#| label: meta-rmse-lognormal-timepoint-interaction

model_path <- here::here("output", "analysis", "fit_rmse_lognormal_timepoint_interaction.rds")

if (refit_models || !file.exists(model_path)) {
  fit_rmse_lognormal_timepoint_interaction <- brm(
    mean_rmse ~ 1 + model * n_valid_person + (1 | dataset_id) + (1 | unique_id),
    data = df_meta,
    family = lognormal(),
    chains = 4,
    cores = 4,
    iter = 3000,
    warmup = 1000,
    seed = seed
  )

  saveRDS(fit_rmse_lognormal_timepoint_interaction, model_path)
} else {
  fit_rmse_lognormal_timepoint_interaction <- readRDS(model_path)
}
```

# Session Info

```{r}
#| label: session-info
pander::pander(sessionInfo())
```
