---
title: "VAR Predictability Benchmark"
subtitle: "Results"
author: 
 - name: Björn S. Siepe
   orcid: 0000-0002-9558-4648
   affiliations: University of Marburg
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    number-sections: true
    theme: cosmo
    code-fold: true
    code-tools: true
    code-summary: "Show the code"
    fig-width: 7
    fig-height: 4.5
    fig-align: "center"
    embed-resources: true
execute:
  message: false
  warning: false
  eval: true
---

<!--
Read-only report. Reads cached results from output/ and never refits.
Primary outcome: person-level RMSE on [0,1]-rescaled items.
Secondary: R². Figures written to figures/.
-->

# Preparation

```{r setup}
#| include: false
library(here)
library(dplyr)
library(ggplot2)
library(tidyr)

source(here("scripts", "00_plotting_functions.R"))

combined <- readRDS(here("output", "meta", "combined.rds"))
metrics <- readRDS(here("output", "analysis", "metrics.rds"))
dataset_meta <- readRDS(here("output", "analysis", "dataset_meta.rds"))

model_order <- c("mean", "locf", "trend", "ri", "ar", "var", "ml_ar", "ml_var")
model_labels <- c("Mean", "LOCF", "Trend", "RI", "AR", "VAR", "mlAR", "mlVAR")

dat_oos <- combined |>
  filter(set == "oos", is.finite(rmse), rmse > 0) |>
  mutate(model = factor(model, levels = model_order, labels = model_labels))
```

# Descriptives

## Sample overview

`n_total_person` counts rows present in the raw data and may underestimate the true number of scheduled beeps if absent responses were never recorded (implicit missingness). It is used as a proxy for series length, not as a compliance measure.

```{r}
#| label: tbl-sample
# TODO: sample summary table (datasets × persons × items × timepoints)
```

## Exclusions

```{r}
#| label: tbl-exclusions
# TODO: exclusion counts by reason (low_obs, zero_var)
```

## Non-Convergence

```{r}
#| label: nonconvergence
# TODO: tabulate non-convergence counts by model and dataset
```


# Predictive accuracy across the complexity ladder

## Average RMSE across datasets

We first look at the average RMSE across all datasets for each model. The RMSE values are calculated per person and per variable, and we average them across all variables for each person in a dataset.
```{r}
#| label: fig-rmse-ladder
#| fig-cap: "Out-of-sample RMSE by model complexity. Each point is one person."
#| fig-height: 5.5
y_cap <- 0.75

outliers <- dat_oos |>
  filter(rmse > y_cap) |>
  count(model, name = "n_out")

message(sprintf("%d observations (%.1f%%) with RMSE > %.2f beyond cap",
                sum(outliers$n_out), 100 * sum(outliers$n_out) / nrow(dat_oos), y_cap))

ggplot(dat_oos, aes(x = model, y = rmse)) +
  ggdist::stat_halfeye(
    adjust = 1, width = 0.5, justification = -0.2,
    .width = 0, point_colour = NA,
    fill = "#A0B4C8", alpha = 0.7
  ) +
  geom_boxplot(
    width = 0.15, outlier.shape = NA, alpha = 0.5
  ) +
  geom_jitter(
    width = 0.05, alpha = 0.15, size = 0.8, colour = "#2C3E50"
  ) +
  geom_segment(
    data = outliers,
    aes(x = model, xend = model, y = y_cap * 0.92, yend = y_cap),
    arrow = arrow(length = unit(0.15, "cm")),
    colour = "grey30", inherit.aes = FALSE
  ) +
  geom_text(
    data = outliers,
    aes(x = model, y = y_cap * 0.88, label = paste0("n=", n_out, "outliers")),
    size = 3, colour = "grey30", inherit.aes = FALSE
  ) +
  coord_cartesian(ylim = c(0, y_cap), expand = FALSE) +
  labs(x = NULL, y = "Out-of-sample RMSE [0, 1]") +
  theme_bs()
```

## In-sample vs. out-of-sample RMSE

```{r}
#| label: fig-rmse-gap
#| fig-cap: "In-sample vs. out-of-sample RMSE by model complexity. Points are means across persons. Lines connect the two sets within each model."
#| fig-height: 5

dat_both <- combined |>
  filter(is.finite(rmse), rmse > 0) |>
  mutate(model = factor(model, levels = model_order, labels = model_labels),
         set = factor(set, levels = c("in", "oos"), labels = c("In-sample", "OOS"))) |>
  summarise(rmse = mean(rmse, na.rm = TRUE), .by = c(model, set))

ggplot(dat_both, aes(x = set, y = rmse, group = model, colour = model)) +
  geom_line(linewidth = 0.7, alpha = 0.8) +
  geom_point(size = 2.5) +
  coord_cartesian(ylim = c(0, 0.3), expand = FALSE) +
  scale_colour_manual(values = model_palette) +
  labs(x = NULL, y = "Mean RMSE [0, 1]", colour = NULL) +
  theme_bs()
```

We investigate the somewhat surprising result that the in-sample RMSE is not always lower than the out-of-sample RMSE, particularly for LOCF where we would not expect that: 

```{r}
#| label: rmse-oos-locf

# per-dataset IS-OOS gap for mean (baseline distributional difference) and locf.
# locf_specific = LOCF gap - Mean gap: the part not explained by OOS time points
# having generally lower variance.
locf_gap <- combined |>
  filter(model %in% c("mean", "locf"), set %in% c("in", "oos"),
         is.finite(rmse), rmse > 0) |>
  summarise(rmse = mean(rmse, na.rm = TRUE), .by = c(dataset_id, model, set)) |>
  pivot_wider(names_from = set, values_from = rmse) |>
  mutate(gap = `in` - oos) |>
  select(dataset_id, model, gap) |>
  pivot_wider(names_from = model, values_from = gap) |>
  mutate(locf_specific = locf - mean) |>
  left_join(select(dataset_meta, dataset_id, first_author, year), by = "dataset_id")

locf_gap |>
  select(first_author, year, mean_gap = mean, locf_gap = locf, locf_specific) |>
  arrange(desc(locf_specific)) |>
  knitr::kable(digits = 3,
               col.names = c("Author", "Year", "Mean gap", "LOCF gap", "LOCF-specific"),
               caption = "IS - OOS RMSE gap per dataset. 'LOCF-specific' removes the baseline distributional gap (Mean gap) from the LOCF gap. Consistently positive values suggest a LOCF-specific mechanism beyond distributional IS/OOS differences.")
```


## Person-specific vs. multilevel models

```{r}
#| label: fig-shrinkage
#| fig-cap: "Person-level OOS RMSE: person-specific vs. multilevel models. Points below the diagonal indicate shrinkage improved predictions."
#| fig-height: 5

dat_shrink <- combined |>
  filter(set == "oos", is.finite(rmse), rmse > 0) |>
  select(dataset_id, id, model, rmse) |>
  pivot_wider(names_from = model, values_from = rmse) |>
  pivot_longer(
    cols = c(ml_ar, ml_var),
    names_to = "ml_model", values_to = "rmse_ml"
  ) |>
  mutate(
    person_model = if_else(ml_model == "ml_ar", ar, var),
    comparison = factor(ml_model,
                        levels = c("ml_ar", "ml_var"),
                        labels = c("AR vs. mlAR", "VAR vs. mlVAR"))
  ) |>
  filter(is.finite(rmse_ml), is.finite(person_model))

xy_cap <- 0.75

ggplot(dat_shrink, aes(x = person_model, y = rmse_ml)) +
  annotate("polygon",
           x = c(0, xy_cap, xy_cap), y = c(0, 0, xy_cap),
           fill = "#6BAED6", alpha = 0.07) +
  annotate("polygon",
           x = c(0, 0, xy_cap), y = c(0, xy_cap, xy_cap),
           fill = "#FC8D59", alpha = 0.07) +
  annotate("text", x = xy_cap * 0.75, y = xy_cap * 0.18,
           label = "Multilevel\nbetter", colour = "#6BAED6",
           size = 3.5, hjust = 0.5) +
  annotate("text", x = xy_cap * 0.18, y = xy_cap * 0.75,
           label = "Person-specific\nbetter", colour = "#FC8D59",
           size = 3.5, hjust = 0.5) +
  geom_abline(slope = 1, intercept = 0, colour = "grey50", linetype = "dashed") +
  geom_point(alpha = 0.2, size = 0.9, colour = "#2C3E50") +
  ggh4x::facet_wrap2(~ comparison, axes = "all") +
  coord_cartesian(xlim = c(0, xy_cap), ylim = c(0, xy_cap), expand = FALSE) +
  labs(x = "Person-specific RMSE", y = "Multilevel RMSE") +
  theme_bs()
```



## Incremental RMSE reduction

```{r}
#| label: fig-incremental
#| fig-cap: "Incremental OOS RMSE reduction at each step of the complexity ladder, with posterior credible intervals. Requires fitted brms model."
#| eval: false
# TODO: successive-difference contrasts from brms model (04_meta_regression.qmd)
# posterior means + 80/95% intervals per complexity step
```

## Heterogeneity across datasets

```{r}
#| label: fig-heterogeneity
#| fig-cap: "OOS RMSE by model and dataset. Each panel is one dataset; points are persons."
#| fig-height: 10

ggplot(dat_oos, aes(y = model, x = rmse)) +
  geom_boxplot(width = 0.5, outlier.shape = NA, alpha = 0.5) +
  geom_jitter(height = 0.15, alpha = 0.15, size = 0.6, colour = "#2C3E50") +
  ggh4x::facet_wrap2(~ dataset_id, axes = "all") +
  coord_cartesian(xlim = c(0, y_cap), expand = FALSE) +
  labs(y = NULL, x = "OOS RMSE [0, 1]") +
  theme_bs(base_size = 11)
```

## Model dominance

```{r}
#| label: fig-dominance
#| fig-cap: "Pairwise model dominance: Proportion of persons where the row model has lower OOS RMSE than the column model. Values above 0.5 indicate the row model wins more often."
#| fig-height: 5.5

model_pairs <- expand.grid(model_a = model_labels, model_b = model_labels,
                           stringsAsFactors = FALSE)

dat_wide <- dat_oos |>
  select(dataset_id, id, model, rmse) |>
  pivot_wider(names_from = model, values_from = rmse)

dominance <- model_pairs |>
  rowwise() |>
  mutate(
    prop = if_else(model_a == model_b, NA_real_,
                   mean(dat_wide[[model_a]] < dat_wide[[model_b]], na.rm = TRUE))
  ) |>
  ungroup() |>
  mutate(
    model_a = factor(model_a, levels = model_labels),
    model_b = factor(model_b, levels = rev(model_labels))
  )

ggplot(dominance, aes(x = model_a, y = model_b, fill = prop)) +
  geom_tile() +
  geom_text(aes(label = round(prop, 2)), size = 3.5, colour = "grey10") +
  scale_fill_distiller(palette = "RdBu", limits = c(0, 1), direction = 1,
                       na.value = "grey95",
                       name = "P(column model better)") +
  labs(x = NULL, y = NULL) +
  scale_x_discrete(position = "top") +
  theme_bs() +
  theme(legend.position = "right")
```

# Model-Based Estimates of Predictive Accuracy

