---
title: "VAR Predictability Benchmark"
subtitle: "Results"
author: 
 - name: Björn S. Siepe
   orcid: 0000-0002-9558-4648
   affiliations: University of Marburg
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    number-sections: true
    theme: cosmo
    code-fold: true
    code-tools: true
    code-summary: "Show the code"
    fig-width: 7
    fig-height: 4.5
    fig-align: "center"
    embed-resources: true
execute:
  message: false
  warning: false
  eval: true
---

<!--
Read-only report. Reads cached results from output/ and never refits.
Primary outcome: person-level RMSE on [0,1]-rescaled items.
Secondary: R². Figures written to figures/.
-->

# Preparation

```{r setup}
#| include: false
library(here)
library(dplyr)
library(ggplot2)
library(tidyr)
library(ggh4x)
library(ggdist)
library(tidybayes)

source(here("scripts", "00_plotting_functions.R"))

combined <- readRDS(here("output", "meta", "combined.rds"))
metrics <- readRDS(here("output", "analysis", "metrics.rds"))
dataset_meta <- readRDS(here("output", "analysis", "dataset_meta.rds"))
features <- readRDS(here("output", "analysis", "features.rds"))

model_order <- c("mean", "locf", "trend", "ri", "ar", "var", "ml_ar", "ml_var")
model_labels <- c("Mean", "LOCF", "Trend", "RI", "AR", "VAR", "mlAR", "mlVAR")

dat_oos <- combined |>
  filter(set == "oos", is.finite(rmse), rmse > 0) |>
  mutate(model = factor(model, levels = model_order, labels = model_labels))
```

# Descriptives

## Sample overview

`n_total_person` counts rows present in the raw data and may underestimate the true number of scheduled beeps if absent responses were never recorded (implicit missingness). It is used as a proxy for series length, not as a compliance measure.

```{r}
#| label: tbl-sample
# TODO: sample summary table (datasets × persons × items × timepoints)
```

## Exclusions

```{r}
#| label: tbl-exclusions
# TODO: exclusion counts by reason (low_obs, zero_var)
```

## Non-Convergence

```{r}
#| label: nonconvergence
# TODO: tabulate non-convergence counts by model and dataset
```


# Predictive accuracy across the complexity ladder

## Average RMSE across datasets

We first look at the average RMSE across all datasets for each model. The RMSE values are calculated per person and per variable, and we average them across all variables for each person in a dataset.
```{r}
#| label: fig-rmse-ladder
#| fig-cap: "Out-of-sample RMSE by model complexity. Each point is one person."
#| fig-height: 5.5
y_cap <- 0.75

outliers <- dat_oos |>
  filter(rmse > y_cap) |>
  count(model, name = "n_out")

message(sprintf("%d observations (%.1f%%) with RMSE > %.2f beyond cap",
                sum(outliers$n_out), 100 * sum(outliers$n_out) / nrow(dat_oos), y_cap))

ggplot(dat_oos, aes(x = model, y = rmse)) +
  ggdist::stat_halfeye(
    adjust = 1, width = 0.5, justification = -0.2,
    .width = 0, point_colour = NA,
    fill = "#A0B4C8", alpha = 0.7
  ) +
  geom_boxplot(
    width = 0.15, outlier.shape = NA, alpha = 0.5
  ) +
  geom_jitter(
    width = 0.05, alpha = 0.15, size = 0.8, colour = "#2C3E50"
  ) +
  geom_segment(
    data = outliers,
    aes(x = model, xend = model, y = y_cap * 0.92, yend = y_cap),
    arrow = arrow(length = unit(0.15, "cm")),
    colour = "grey30", inherit.aes = FALSE
  ) +
  geom_text(
    data = outliers,
    aes(x = model, y = y_cap * 0.88, label = paste0("n=", n_out, "outliers")),
    size = 3, colour = "grey30", inherit.aes = FALSE
  ) +
  coord_cartesian(ylim = c(0, y_cap), expand = FALSE) +
  labs(x = NULL, y = "Out-of-sample RMSE [0, 1]") +
  theme_bs()
```

## In-sample vs. out-of-sample RMSE

We here compute the in-sample performance by fitting each model on the whole time series, without a test-set split.

```{r}
#| label: fig-rmse-gap
#| fig-cap: "In-sample vs. out-of-sample RMSE by model complexity. Points are means across persons. Lines connect the two sets within each model."
#| fig-height: 5

dat_both <- combined |>
  filter(is.finite(rmse), rmse > 0) |>
  mutate(model = factor(model, levels = model_order, labels = model_labels),
         set = factor(set, levels = c("in", "oos"), labels = c("In-sample", "OOS"))) |>
  summarise(rmse = mean(rmse, na.rm = TRUE), .by = c(model, set))

ggplot(dat_both, aes(x = set, y = rmse, group = model, colour = model)) +
  geom_line(linewidth = 0.7, alpha = 0.8) +
  geom_point(size = 2.5) +
  coord_cartesian(ylim = c(0, 0.3), expand = FALSE) +
  scale_colour_manual(values = model_palette) +
  labs(x = NULL, y = "Mean RMSE [0, 1]", colour = NULL) +
  theme_bs()
```

We investigate the somewhat surprising result that the in-sample RMSE is not always lower than the out-of-sample RMSE, particularly for LOCF where we would not expect that: 

```{r}
#| label: rmse-oos-locf
# per-dataset IS-OOS gap for mean (baseline distributional difference) and locf.
# locf_specific = LOCF gap - Mean gap: the part not explained by OOS time points
# having generally lower variance.
locf_gap <- combined |>
  filter(model %in% c("mean", "locf"), set %in% c("in", "oos"),
         is.finite(rmse), rmse > 0) |>
  summarise(rmse = mean(rmse, na.rm = TRUE), .by = c(dataset_id, model, set)) |>
  pivot_wider(names_from = set, values_from = rmse) |>
  mutate(gap = `in` - oos) |>
  select(dataset_id, model, gap) |>
  pivot_wider(names_from = model, values_from = gap) |>
  mutate(locf_specific = locf - mean) |>
  left_join(dataset_meta |> select(dataset_id, first_author, year), by = "dataset_id")

locf_gap |>
  select(first_author, year, mean_gap = mean, locf_gap = locf, locf_specific) |>
  arrange(desc(locf_specific)) |>
  knitr::kable(digits = 3,
               col.names = c("Author", "Year", "Mean gap", "LOCF gap", "LOCF-specific"),
               caption = "IS - OOS RMSE gap per dataset. 'LOCF-specific' removes the baseline distributional gap (Mean gap) from the LOCF gap. Consistently positive values suggest a LOCF-specific mechanism beyond distributional IS/OOS differences.")
```




## Person-specific vs. multilevel models

```{r}
#| label: fig-shrinkage
#| fig-cap: "Person-level OOS RMSE: person-specific vs. multilevel models. Points below the diagonal indicate shrinkage improved predictions."
#| fig-height: 5

dat_shrink <- combined |>
  filter(set == "oos", is.finite(rmse), rmse > 0) |>
  select(dataset_id, id, model, rmse) |>
  pivot_wider(names_from = model, values_from = rmse) |>
  pivot_longer(
    cols = c(ml_ar, ml_var),
    names_to = "ml_model", values_to = "rmse_ml"
  ) |>
  mutate(
    person_model = if_else(ml_model == "ml_ar", ar, var),
    comparison = factor(ml_model,
                        levels = c("ml_ar", "ml_var"),
                        labels = c("AR vs. mlAR", "VAR vs. mlVAR"))
  ) |>
  filter(is.finite(rmse_ml), is.finite(person_model))

xy_cap <- 0.75

ggplot(dat_shrink, aes(x = person_model, y = rmse_ml)) +
  annotate("polygon",
           x = c(0, xy_cap, xy_cap), y = c(0, 0, xy_cap),
           fill = "#6BAED6", alpha = 0.07) +
  annotate("polygon",
           x = c(0, 0, xy_cap), y = c(0, xy_cap, xy_cap),
           fill = "#FC8D59", alpha = 0.07) +
  annotate("text", x = xy_cap * 0.75, y = xy_cap * 0.18,
           label = "Multilevel\nbetter", colour = "#6BAED6",
           size = 3.5, hjust = 0.5) +
  annotate("text", x = xy_cap * 0.18, y = xy_cap * 0.75,
           label = "Person-specific\nbetter", colour = "#FC8D59",
           size = 3.5, hjust = 0.5) +
  geom_abline(slope = 1, intercept = 0, colour = "grey50", linetype = "dashed") +
  geom_point(alpha = 0.2, size = 0.9, colour = "#2C3E50") +
  ggh4x::facet_wrap2(~ comparison, axes = "all") +
  coord_cartesian(xlim = c(0, xy_cap), ylim = c(0, xy_cap), expand = FALSE) +
  labs(x = "Person-specific RMSE", y = "Multilevel RMSE") +
  theme_bs()
```



## Incremental RMSE reduction

```{r}
#| label: fig-incremental
#| fig-cap: "Incremental OOS RMSE reduction at each step of the complexity ladder, with posterior credible intervals. Requires fitted brms model."
#| eval: false
# TODO: successive-difference contrasts from brms model (04_meta_regression.qmd)
# posterior means + 80/95% intervals per complexity step
```

## Heterogeneity across datasets

```{r}
#| label: fig-heterogeneity
#| fig-cap: "OOS RMSE by model and dataset. Each panel is one dataset; points are persons."
#| fig-height: 10

ggplot(dat_oos, aes(y = model, x = rmse)) +
  geom_boxplot(width = 0.5, outlier.shape = NA, alpha = 0.5) +
  geom_jitter(height = 0.15, alpha = 0.15, size = 0.6, colour = "#2C3E50") +
  ggh4x::facet_wrap2(~ dataset_id, axes = "all") +
  coord_cartesian(xlim = c(0, y_cap), expand = FALSE) +
  labs(y = NULL, x = "OOS RMSE [0, 1]") +
  theme_bs(base_size = 11)
```

## Model dominance

```{r}
#| label: fig-dominance
#| fig-cap: "Pairwise model dominance: Proportion of persons where the row model has lower OOS RMSE than the column model. Values above 0.5 indicate the row model wins more often."
#| fig-height: 5.5

model_pairs <- expand.grid(model_a = model_labels, model_b = model_labels,
                           stringsAsFactors = FALSE)

dat_wide <- dat_oos |>
  select(dataset_id, id, model, rmse) |>
  pivot_wider(names_from = model, values_from = rmse)

dominance <- model_pairs |>
  rowwise() |>
  mutate(
    prop = if_else(model_a == model_b, NA_real_,
                   mean(dat_wide[[model_a]] < dat_wide[[model_b]], na.rm = TRUE))
  ) |>
  ungroup() |>
  mutate(
    model_a = factor(model_a, levels = model_labels),
    model_b = factor(model_b, levels = rev(model_labels))
  )

ggplot(dominance, aes(x = model_a, y = model_b, fill = prop)) +
  geom_tile() +
  geom_text(aes(label = round(prop, 2)), size = 3.5, colour = "grey10") +
  scale_fill_distiller(palette = "RdBu", limits = c(0, 1), direction = 1,
                       na.value = "grey95",
                       name = "P(column model better)") +
  labs(x = NULL, y = NULL) +
  scale_x_discrete(position = "top") +
  theme_bs() +
  theme(legend.position = "right")
```

## Decomposition of predictive gains

Here, we display total model improvement from the mean model to the most complex model (multilevel VAR) and decompose it into three components: autoregression, cross-dynamics, and pooling. We also show two different paths of improvement: first adding cross-dynamics and then pooling, or first pooling and then adding cross-dynamics.

```{r}
#| label: fig-decomposition
#| fig-cap: "Decomposition of the change in OOS RMSE from the mean model to mlVAR, along two paths."
#| fig-height: 9
#| fig-width: 11

decomp_models <- c("mean", "ar", "var", "ml_ar", "ml_var")
component_levels <- c("Autoregression", "Cross-dynamics", "Pooling", "Total")
path_levels <- c("A: cross-lags,\nthen pooling", "B: pooling,\nthen cross-lags")
component_palette <- setNames(c(MetBrewer::met.brewer("Hiroshige")[c(9, 2, 7)], "grey45"),
                              component_levels)

rmse_wide <- combined |>
  filter(set == "oos", model %in% decomp_models) |>
  mutate(rmse = if_else(is.finite(rmse) & rmse > 0, rmse, NA_real_)) |>
  select(dataset_id, first_author, year, id, model, rmse) |>
  pivot_wider(names_from = model, values_from = rmse)

# persons need all five models for the components to add up
n_incomplete <- sum(!complete.cases(rmse_wide[decomp_models]))
rmse_wide <- filter(rmse_wide, if_all(all_of(decomp_models), is.finite))

# change in RMSE per step, negative values mean lower RMSE after the step
id_cols <- c("dataset_id", "first_author", "year", "id", "path")
deltas <- bind_rows(
  rmse_wide |>
    mutate(path = path_levels[1],
           Autoregression = ar - mean, `Cross-dynamics` = var - ar,
           Pooling = ml_var - var, Total = ml_var - mean),
  rmse_wide |>
    mutate(path = path_levels[2],
           Autoregression = ar - mean, `Cross-dynamics` = ml_var - ml_ar,
           Pooling = ml_ar - ar, Total = ml_var - mean)
) |>
  select(all_of(id_cols), all_of(component_levels)) |>
  pivot_longer(all_of(component_levels), names_to = "component", values_to = "delta")

delta_ds <- deltas |>
  summarise(est = mean(delta), .by = c(dataset_id, first_author, year, path, component))

# datasets are weighted equally
delta_all <- delta_ds |>
  summarise(est = mean(est), .by = c(path, component)) |>
  mutate(label = "All datasets")

ds_labels <- delta_ds |>
  distinct(dataset_id, first_author, year) |>
  mutate(label = paste0(first_author, " (", year, ")"),
         label = if_else(duplicated(label) | duplicated(label, fromLast = TRUE),
                         paste0(label, " [", dataset_id, "]"), label))

# largest total improvement at the top
label_order <- delta_ds |>
  filter(component == "Total", path == path_levels[1]) |>
  left_join(ds_labels, by = c("dataset_id", "first_author", "year")) |>
  arrange(desc(est)) |>
  pull(label)

# symmetric x-range shared by all panels. bars beyond it are cut short and labelled with their value
x_lim <- 0.1

plot_decomp <- delta_ds |>
  left_join(ds_labels, by = c("dataset_id", "first_author", "year")) |>
  bind_rows(delta_all) |>
  mutate(capped = abs(est) > x_lim,
         bar_end = if_else(capped, sign(est) * x_lim * 0.97, est),
         label = factor(label, levels = c("All datasets", label_order)),
         component = factor(component, levels = component_levels),
         path = factor(path, levels = path_levels))

ggplot(plot_decomp, aes(x = bar_end, y = label, fill = component)) +
  geom_vline(xintercept = 0, colour = "grey50", linewidth = 0.3) +
  geom_col(width = 0.7) +
  geom_text(data = filter(plot_decomp, capped),
            aes(x = sign(est) * x_lim, label = sprintf("%.2f", est),
                hjust = if_else(est > 0, 1, 0)),
            size = 2.8, colour = "grey20") +
  geom_hline(yintercept = 1.5, colour = "grey70", linewidth = 0.3) +
  scale_fill_manual(values = component_palette, guide = "none") +
  ggh4x::facet_grid2(path ~ component, axes = "all", remove_labels = "y") +
  coord_cartesian(xlim = c(-x_lim, x_lim), expand = FALSE) +
  labs(x = "Change in OOS RMSE [0, 1] (< 0 = better)", y = NULL) +
  theme_bs(base_size = 11) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_line(colour = "#E0E0E0"))
```

## Value of pooling by series length

```{r}
#| label: fig-pooling-length
#| fig-cap: "Change in OOS RMSE from pooling, by the person's training series length."
#| fig-height: 5
#| fig-width: 9

y_lim <- c(-0.1, 0.05)

pooling <- combined |>
  filter(set == "oos", model %in% c("ar", "ml_ar", "var", "ml_var"),
         is.finite(rmse), rmse > 0) |>
  select(dataset_id, id, model, rmse) |>
  pivot_wider(names_from = model, values_from = rmse) |>
  mutate(`AR → mlAR` = ml_ar - ar, `VAR → mlVAR` = ml_var - var) |>
  select(dataset_id, id, `AR → mlAR`, `VAR → mlVAR`) |>
  pivot_longer(-c(dataset_id, id), names_to = "comparison", values_to = "delta") |>
  filter(is.finite(delta)) |>
  left_join(select(features$person, dataset_id, id, n_train_first),
            by = c("dataset_id", "id"))

stopifnot(!anyNA(pooling$n_train_first))

beyond <- pooling |>
  summarise(n_below = sum(delta < y_lim[1]), n_above = sum(delta > y_lim[2]),
            .by = comparison) |>
  mutate(label = sprintf("%d above, %d below range", n_above, n_below))

# robust loess downweights persons whose person-specific model breaks down
ggplot(pooling, aes(x = n_train_first, y = delta)) +
  geom_hline(yintercept = 0, colour = "grey50", linewidth = 0.3) +
  geom_point(alpha = 0.15, size = 0.8, colour = "#2C3E50") +
  geom_smooth(method = "loess", formula = y ~ x, se = TRUE,
              method.args = list(family = "symmetric"),
              colour = component_palette[["Pooling"]], linewidth = 1) +
  # draw one line per dataset to show the within-dataset trend
  geom_smooth(aes(group = dataset_id), method = "lm", se = FALSE,
              colour = scales::alpha("grey50", 0.6), linewidth = 0.5) +
  geom_text(data = beyond, aes(x = -Inf, y = Inf, label = label),
            inherit.aes = FALSE, hjust = -0.05, vjust = 1.5, size = 3, colour = "grey30") +
  scale_x_log10() +
  ggh4x::facet_wrap2(~ comparison, axes = "all") +
  coord_cartesian(ylim = y_lim) +
  labs(x = "Training series length (first fold, log scale)",
       y = "Change in OOS RMSE [0, 1] (< 0 = pooling better)") +
  theme_bs()
```

## Value of cross-dynamics by data per parameter

```{r}
#| label: fig-cross-dynamics
#| fig-cap: "Change in OOS RMSE from adding cross-lagged effects, by the number of training observations per VAR coefficient."
#| fig-height: 5
#| fig-width: 9

y_lim_cross <- c(-0.05, 0.1)

cross_dyn <- combined |>
  filter(set == "oos", model %in% c("ar", "var", "ml_ar", "ml_var"),
         is.finite(rmse), rmse > 0) |>
  select(dataset_id, id, model, rmse) |>
  pivot_wider(names_from = model, values_from = rmse) |>
  mutate(`AR → VAR` = var - ar, `mlAR → mlVAR` = ml_var - ml_ar) |>
  select(dataset_id, id, `AR → VAR`, `mlAR → mlVAR`) |>
  pivot_longer(-c(dataset_id, id), names_to = "comparison", values_to = "delta") |>
  filter(is.finite(delta)) |>
  left_join(select(features$person, dataset_id, id, obs_per_var_eq_par),
            by = c("dataset_id", "id"))

stopifnot(!anyNA(cross_dyn$obs_per_var_eq_par))

beyond_cross <- cross_dyn |>
  summarise(n_below = sum(delta < y_lim_cross[1]), n_above = sum(delta > y_lim_cross[2]),
            .by = comparison) |>
  mutate(label = sprintf("%d above, %d below range", n_above, n_below))

ggplot(cross_dyn, aes(x = obs_per_var_eq_par, y = delta)) +
  geom_hline(yintercept = 0, colour = "grey50", linewidth = 0.3) +
  geom_point(alpha = 0.15, size = 0.8, colour = "#2C3E50") +
  geom_smooth(method = "loess", formula = y ~ x, se = TRUE,
              method.args = list(family = "symmetric"),
              colour = component_palette[["Cross-dynamics"]], linewidth = 1) +
  # draw one line per dataset to show the within-dataset trend
  geom_smooth(aes(group = dataset_id), method = "lm", se = FALSE,
              colour = scales::alpha("grey50", 0.6), linewidth = 0.5) +
  geom_text(data = beyond_cross, aes(x = -Inf, y = Inf, label = label),
            inherit.aes = FALSE, hjust = -0.05, vjust = 1.5, size = 3, colour = "grey30") +
  scale_x_log10() +
  ggh4x::facet_wrap2(~ comparison, axes = "all") +
  coord_cartesian(ylim = y_lim_cross) +
  labs(x = "Training observations per VAR coefficient (first fold, log scale)",
       y = "Change in OOS RMSE [0, 1] (< 0 = cross-lags better)") +
  theme_bs()
```

## Time series gallery

We now select a few persons with good, average, and poor predictions by mlVAR relative to the mean model, and show one typical item per person. We also show the first OOS origin as a dashed line. We here define "skill" as the relative improvement of mlVAR over the mean model, i.e., 1 - RMSE_mlVAR / RMSE_mean. A skill of 0 means that mlVAR performs equally well as the mean model, a skill of 0.5 means that mlVAR reduces the RMSE by 50% relative to the mean model, and a negative skill means that mlVAR performs worse than the mean model.


```{r}
#| label: fig-gallery
#| fig-cap: "Observed series and OOS predictions for the persons with the highest, median and lowest mlVAR skill relative to the mean model. One typical item per person. The dashed line marks the first OOS origin."
#| fig-height: 9
#| fig-width: 11

# person-level skill of mlVAR relative to the mean model
person_skill <- combined |>
  filter(set == "oos", model %in% c("mean", "ml_var"), is.finite(rmse), rmse > 0) |>
  select(dataset_id, first_author, year, id, model, rmse) |>
  pivot_wider(names_from = model, values_from = rmse, names_prefix = "rmse_") |>
  filter(is.finite(rmse_mean), is.finite(rmse_ml_var)) |>
  mutate(skill = 1 - rmse_ml_var / rmse_mean) |>
  arrange(skill, dataset_id, id)

median_skill <- median(person_skill$skill)
selected <- bind_rows(
  slice_max(person_skill, skill, n = 3, with_ties = FALSE) |> mutate(group = "Best"),
  slice_min(person_skill, abs(skill - median_skill), n = 3, with_ties = FALSE) |>
    mutate(group = "Median"),
  slice_min(person_skill, skill, n = 3, with_ties = FALSE) |> mutate(group = "Worst")
) |>
  mutate(rank = row_number(), .by = group)

# per person, the item whose skill is closest to that person's median item skill
item_skill <- features$item_error |>
  filter(model %in% c("mean", "ml_var")) |>
  select(dataset_id, id, variable, model, rmse) |>
  pivot_wider(names_from = model, values_from = rmse, names_prefix = "rmse_") |>
  filter(rmse_mean > 0) |>
  mutate(item_skill = 1 - rmse_ml_var / rmse_mean)

gallery_items <- selected |>
  select(dataset_id, first_author, year, id, skill, group, rank) |>
  inner_join(item_skill, by = c("dataset_id", "id")) |>
  slice_min(abs(item_skill - median(item_skill)), n = 1, with_ties = FALSE,
            by = c(dataset_id, id)) |>
  mutate(group = factor(group, levels = c("Best", "Median", "Worst"))) |>
  arrange(group, rank) |>
  mutate(panel = sprintf("%s %d, skill %.2f\n%s (%s), %s",
                         group, rank, skill, first_author, year, variable),
         panel = factor(panel, levels = panel))

gallery_keys <- select(gallery_items, dataset_id, id, variable, panel)

gallery_series <- features$series |>
  inner_join(gallery_keys, by = c("dataset_id", "id", "variable"))

gallery_preds <- features$oos |>
  filter(model %in% c("mean", "ml_var")) |>
  inner_join(gallery_keys, by = c("dataset_id", "id", "variable")) |>
  mutate(model = factor(model, levels = c("mean", "ml_var"), labels = c("Mean", "mlVAR")))

first_origin <- summarise(gallery_preds, t0 = min(t), .by = panel)

ggplot(gallery_series, aes(x = t, y = y)) +
  geom_vline(data = first_origin, aes(xintercept = t0 - 0.5),
             linetype = "dashed", colour = "grey50", linewidth = 0.3) +
  geom_line(colour = "grey60", linewidth = 0.3) +
  geom_point(colour = "grey30", size = 0.5) +
  geom_line(data = gallery_preds, aes(y = yhat, colour = model), linewidth = 0.6) +
  geom_point(data = gallery_preds, aes(y = yhat, colour = model), size = 1) +
  scale_colour_manual(values = model_palette, name = NULL) +
  ggh4x::facet_wrap2(~ panel, ncol = 3, scales = "free_x", axes = "all") +
  coord_cartesian(ylim = c(0, 1)) +
  labs(x = "Measurement occasion", y = "Response [0, 1]") +
  theme_bs(base_size = 11)
```


# Model-Based Estimates of Predictive Accuracy


## M1 (Model using only model as predictor)

We estimated a log-normal meta-regression model with only the model as a predictor. We can now use the outputs of this model to create a forest plot-like visualization of the predictive accuracy of each model across datasets. 

TODO I'm not sure if that is the best way to visualize this - the dataset specific effects might not really come through here. 

```{r}
#| label: fig-caterpillar-m1
#| fig-cap: "Dataset-level predicted RMSE by model. Points are posterior means; thick bars are 80% CIs, thin lines are 95% CIs."
#| fig-height: 10
#| fig-width: 9

fit_rmse_lognormal <- readRDS(here::here("output", "analysis", "fit_rmse_lognormal.rds"))

newdata <- expand.grid(
  model = levels(fit_rmse_lognormal$data$model),
  dataset_id = unique(fit_rmse_lognormal$data$dataset_id),
  unique_id = NA,
  stringsAsFactors = FALSE
)

preds <- newdata |>
  add_epred_draws(fit_rmse_lognormal, re_formula = ~ (1 | dataset_id) + (1 | dataset_id:model),
                  # for the dataset-level predictions, we do not include the person-level random effect
                  allow_new_levels = TRUE) |>
  ungroup() |>
  mutate(model = factor(model, levels = model_order, labels = model_labels))

preds <- preds |>
  left_join(dataset_meta |> select(dataset_id, first_author, year), by = "dataset_id") |>
  mutate(dataset_label = paste0(first_author, " (", year, ")"))

ggplot(preds, aes(x = .epred, y = reorder(dataset_label, .epred),
                  colour = model)) +
  stat_pointinterval(position = position_dodge(width = 0.7),
                     .width = c(0.80, 0.95),
                     point_size = 1.5, interval_size_range = c(0.4, 1.2)) +
  scale_colour_manual(values = model_palette) +
  coord_cartesian(xlim = c(0, 0.5), expand = FALSE) +
  labs(x = "Predicted RMSE [0, 1]", y = NULL, colour = NULL) +
  theme_bs() +
  theme(legend.position = "bottom",
        panel.grid.major.y = element_blank())+
  geom_hline(yintercept = seq(1.5, length(unique(preds$dataset_label)) - 0.5, 1),
           colour = "grey90", linewidth = 0.3) +
  guides(colour = guide_legend(nrow = 1, byrow = TRUE))
```