---
title: "VAR Predictability Benchmark"
subtitle: "Results"
author: 
 - name: Björn S. Siepe
   orcid: 0000-0002-9558-4648
   affiliations: University of Marburg
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    number-sections: true
    theme: cosmo
    code-fold: true
    code-tools: true
    code-summary: "Show the code"
    fig-width: 7
    fig-height: 4.5
    fig-align: "center"
    embed-resources: true
execute:
  message: false
  warning: false
  eval: true
---

<!--
Read-only report. Reads cached results from output/ and never refits.
Primary outcome: person-level RMSE on [0,1]-rescaled items.
Secondary: R². Figures written to figures/.
-->

# Preparation

```{r setup}
#| include: false
library(here)
library(dplyr)
library(ggplot2)
library(tidyr)
library(ggh4x)
library(ggdist)
library(tidybayes)

source(here("scripts", "00_plotting_functions.R"))

combined <- readRDS(here("output", "meta", "combined.rds"))
metrics <- readRDS(here("output", "analysis", "metrics.rds"))
dataset_meta <- readRDS(here("output", "analysis", "dataset_meta.rds"))
features <- readRDS(here("output", "analysis", "features.rds"))
fit_issues <- readRDS(here("output", "analysis", "fit_issues.rds"))

model_order <- c("mean", "locf", "trend", "ri", "ar", "var", "ml_ar", "ml_var")
model_labels <- c("Mean", "LOCF", "Trend", "RI", "AR", "VAR", "mlAR", "mlVAR")

dat_oos <- combined |>
  filter(set == "oos", is.finite(rmse), rmse > 0) |>
  mutate(model = factor(model, levels = model_order, labels = model_labels))

# plotting params
decomp_models <- c("mean", "ar", "var", "ml_ar", "ml_var")
component_levels <- c("Autoregression", "Cross-dynamics", "Pooling", "Total")
path_levels <- c("A: cross-lags,\nthen pooling", "B: pooling,\nthen cross-lags")
component_palette <- setNames(c(MetBrewer::met.brewer("Hiroshige")[c(9, 2, 7)], "grey45"),
                              component_levels)

```

# Descriptives

## Sample overview

`n_total_person` counts rows present in the raw data and may underestimate the true number of scheduled beeps if absent responses were never recorded (implicit missingness). It is used as a proxy for series length, not as a compliance measure.

```{r}
#| label: tbl-sample
#| tbl-cap: "Sample overview per dataset. Valid observations are timepoints with complete responses and a usable lag."

person_n <- combined |>
  distinct(dataset_id, id, n_valid_person, n_total_person) |>
  summarise(valid_median = median(n_valid_person), valid_min = min(n_valid_person),
            valid_max = max(n_valid_person), rows_median = median(n_total_person),
            .by = dataset_id)

dataset_meta |>
  left_join(person_n, by = "dataset_id") |>
  arrange(first_author, year) |>
  mutate(dataset = paste0(first_author, " (", year, ")"),
         valid = sprintf("%g [%g, %g]", valid_median, valid_min, valid_max)) |>
  select(dataset, n_beeps_per_day, n_persons_kept, n_persons_excluded, n_items,
         valid, rows_median) |>
  knitr::kable(col.names = c("Dataset", "Beeps per day", "Persons", "Excluded", "Items",
                             "Valid obs. per person, median [range]", "Rows per person, median"))
```

## Exclusions

```{r}
#| label: tbl-exclusions
#| tbl-cap: "Persons included and excluded per dataset, by reason."

dataset_meta |>
  arrange(first_author, year) |>
  mutate(dataset = paste0(first_author, " (", year, ")")) |>
  select(dataset, n_persons_kept, n_excluded_low_obs, n_excluded_zero_var) |>
  knitr::kable(col.names = c("Dataset", "Included", "Excluded: too few valid obs.",
                             "Excluded: zero variance"))
```

## Non-Convergence

```{r}
#| label: tbl-nonconvergence
#| tbl-cap: "Estimation issues per dataset. Rank-deficient fits are the share of person-level OLS fits with aliased coefficients (set to 0). Singular fits are the share of lmer fits (one per item and fit step). Convergence and other warnings are counts of warnings."

ols_models <- c("ar", "var")
ml_models <- c("ri", "ml_ar", "ml_var")

# share of affected fits for one issue type, one column per model
issue_share <- function(models, issue, labels, prefix) {
  fit_issues |>
    filter(model %in% models, type == issue) |>
    mutate(share = sprintf("%.0f%%", 100 * n / n_fits),
           model = paste(prefix, factor(model, levels = models, labels = labels))) |>
    select(dataset_id, model, share) |>
    pivot_wider(names_from = model, values_from = share)
}

rank_deficient <- issue_share(ols_models, "rank_deficient", c("AR", "VAR"), "Rank-deficient:")
singular <- issue_share(ml_models, "singular", c("RI", "mlAR", "mlVAR"), "Singular:")

other_issues <- fit_issues |>
  summarise(`Convergence warnings` = sum(n[type == "convergence"]),
            `Other warnings` = sum(n[type == "other"]), .by = dataset_id)

failed <- dataset_meta |>
  mutate(failed = vapply(model_failures, function(f) {
    paste(model_labels[match(names(f)[f], model_order)], collapse = ", ")
  }, character(1))) |>
  select(dataset_id, first_author, year, failed, ml_var_uncorrelated)

failed |>
  left_join(rank_deficient, by = "dataset_id") |>
  left_join(singular, by = "dataset_id") |>
  left_join(other_issues, by = "dataset_id") |>
  arrange(first_author, year) |>
  mutate(Dataset = paste0(first_author, " (", year, ")"),
         `Failed models` = failed,
         `mlVAR uncorrelated RE` = if_else(ml_var_uncorrelated, "yes", "no")) |>
  select(Dataset, starts_with("Rank-deficient"), starts_with("Singular"),
         `Convergence warnings`, `Other warnings`, `Failed models`, `mlVAR uncorrelated RE`) |>
  knitr::kable()
```


# Predictive accuracy across the complexity ladder

## Average RMSE across datasets

We first look at the average RMSE across all datasets for each model. The RMSE values are calculated per person and per variable, and we average them across all variables for each person in a dataset.
```{r}
#| label: fig-rmse-ladder
#| fig-cap: "Out-of-sample RMSE by model complexity. Each point is one person."
#| fig-height: 5.5
y_cap <- 0.75

outliers <- dat_oos |>
  filter(rmse > y_cap) |>
  count(model, name = "n_out")

message(sprintf("%d observations (%.1f%%) with RMSE > %.2f beyond cap",
                sum(outliers$n_out), 100 * sum(outliers$n_out) / nrow(dat_oos), y_cap))

ggplot(dat_oos, aes(x = model, y = rmse)) +
  ggdist::stat_halfeye(
    adjust = 1, width = 0.5, justification = -0.2,
    .width = 0, point_colour = NA,
    fill = "#A0B4C8", alpha = 0.7
  ) +
  geom_boxplot(
    width = 0.15, outlier.shape = NA, alpha = 0.5
  ) +
  geom_jitter(
    width = 0.05, alpha = 0.15, size = 0.8, colour = "#2C3E50"
  ) +
  geom_segment(
    data = outliers,
    aes(x = model, xend = model, y = y_cap * 0.92, yend = y_cap),
    arrow = arrow(length = unit(0.15, "cm")),
    colour = "grey30", inherit.aes = FALSE
  ) +
  geom_text(
    data = outliers,
    aes(x = model, y = y_cap * 0.88, label = paste0("n=", n_out, "outliers")),
    size = 3, colour = "grey30", inherit.aes = FALSE
  ) +
  coord_cartesian(ylim = c(0, y_cap), expand = FALSE) +
  labs(x = NULL, y = "Out-of-sample RMSE [0, 1]") +
  theme_bs()
```

## In-sample vs. out-of-sample RMSE

We here compute the in-sample performance by fitting each model on the whole time series, without a test-set split.

```{r}
#| label: fig-rmse-gap
#| fig-cap: "In-sample vs. out-of-sample RMSE by model complexity. Points are means across persons. Lines connect the two sets within each model."
#| fig-height: 5

dat_both <- combined |>
  filter(is.finite(rmse), rmse > 0) |>
  mutate(model = factor(model, levels = model_order, labels = model_labels),
         set = factor(set, levels = c("in", "oos"), labels = c("In-sample", "OOS"))) |>
  summarise(rmse = mean(rmse, na.rm = TRUE), .by = c(model, set))

ggplot(dat_both, aes(x = set, y = rmse, group = model, colour = model)) +
  geom_line(linewidth = 0.7, alpha = 0.8) +
  geom_point(size = 2.5) +
  coord_cartesian(ylim = c(0, 0.3), expand = FALSE) +
  scale_colour_manual(values = model_palette) +
  labs(x = NULL, y = "Mean RMSE [0, 1]", colour = NULL) +
  theme_bs()
```

We investigate the somewhat surprising result that the in-sample RMSE is not always lower than the out-of-sample RMSE, particularly for LOCF where we would not expect that: 

```{r}
#| label: rmse-oos-locf
# per-dataset IS-OOS gap for mean (baseline distributional difference) and locf.
# locf_specific = LOCF gap - Mean gap: the part not explained by OOS time points
# having generally lower variance.
locf_gap <- combined |>
  filter(model %in% c("mean", "locf"), set %in% c("in", "oos"),
         is.finite(rmse), rmse > 0) |>
  summarise(rmse = mean(rmse, na.rm = TRUE), .by = c(dataset_id, model, set)) |>
  pivot_wider(names_from = set, values_from = rmse) |>
  mutate(gap = `in` - oos) |>
  select(dataset_id, model, gap) |>
  pivot_wider(names_from = model, values_from = gap) |>
  mutate(locf_specific = locf - mean) |>
  left_join(dataset_meta |> select(dataset_id, first_author, year), by = "dataset_id")

locf_gap |>
  select(first_author, year, mean_gap = mean, locf_gap = locf, locf_specific) |>
  arrange(desc(locf_specific)) |>
  knitr::kable(digits = 3,
               col.names = c("Author", "Year", "Mean gap", "LOCF gap", "LOCF-specific"),
               caption = "IS - OOS RMSE gap per dataset. 'LOCF-specific' removes the baseline distributional gap (Mean gap) from the LOCF gap. Consistently positive values suggest a LOCF-specific mechanism beyond distributional IS/OOS differences.")
```

```{r}
#| label: fig-is-oos-barbell
#| fig-cap: "In-sample and OOS RMSE per dataset and model. Points are means across persons. Rows are ordered by the gap of the mean model."
#| fig-height: 10
#| fig-width: 12

barbell_labels <- dataset_meta |>
  distinct(dataset_id, first_author, year) |>
  mutate(dataset = paste0(first_author, " (", year, ")"),
         dataset = if_else(duplicated(dataset) | duplicated(dataset, fromLast = TRUE),
                           paste0(dataset, " [", dataset_id, "]"), dataset))

is_oos <- combined |>
  filter(set %in% c("in", "oos"), is.finite(rmse), rmse > 0) |>
  select(dataset_id, id, model, set, rmse) |>
  pivot_wider(names_from = set, values_from = rmse) |>
  # persons enter only with both sets, so the dataset gap is paired
  filter(is.finite(`in`), is.finite(oos)) |>
  summarise(is = mean(`in`), oos = mean(oos), .by = c(dataset_id, model)) |>
  left_join(barbell_labels, by = "dataset_id") |>
  mutate(direction = if_else(oos > is, "OOS higher", "OOS lower"),
         model = factor(model, levels = model_order, labels = model_labels))

# largest mean-model gap at the top
row_order <- is_oos |>
  filter(model == "Mean") |>
  arrange(oos - is) |>
  pull(dataset)

is_oos <- mutate(is_oos, dataset = factor(dataset, levels = union(row_order, dataset)))

is_oos_points <- is_oos |>
  pivot_longer(c(is, oos), names_to = "set", values_to = "rmse") |>
  mutate(set = factor(set, levels = c("is", "oos"), labels = c("In-sample", "Out-of-sample")))

# points beyond the cap are drawn at the cap and labelled with their value
x_cap <- 0.4

ggplot(is_oos, aes(y = dataset)) +
  geom_segment(aes(x = pmin(is, x_cap), xend = pmin(oos, x_cap), colour = direction),
               linewidth = 0.8) +
  geom_point(data = is_oos_points, aes(x = pmin(rmse, x_cap), fill = set),
             shape = 21, colour = "grey30", size = 1.8, stroke = 0.3) +
  geom_text(data = filter(is_oos_points, rmse > x_cap),
            aes(x = x_cap, label = sprintf("%.2f", rmse)),
            hjust = 1.3, size = 2.5, colour = "grey20") +
  scale_colour_manual(values = c("OOS higher" = "grey70",
                                 "OOS lower" = component_palette[["Cross-dynamics"]]),
                      name = NULL) +
  scale_fill_manual(values = c("In-sample" = "white",
                               "Out-of-sample" = component_palette[["Autoregression"]]),
                    name = NULL) +
  ggh4x::facet_wrap2(~ model, nrow = 2, axes = "all", remove_labels = "y") +
  labs(x = "RMSE [0, 1]", y = NULL) +
  theme_bs(base_size = 11) +
  theme(legend.position = "bottom")
```




## Person-specific vs. multilevel models

```{r}
#| label: fig-shrinkage
#| fig-cap: "Person-level OOS RMSE: person-specific vs. multilevel models. Points below the diagonal indicate shrinkage improved predictions."
#| fig-height: 5

dat_shrink <- combined |>
  filter(set == "oos", is.finite(rmse), rmse > 0) |>
  select(dataset_id, id, model, rmse) |>
  pivot_wider(names_from = model, values_from = rmse) |>
  pivot_longer(
    cols = c(ml_ar, ml_var),
    names_to = "ml_model", values_to = "rmse_ml"
  ) |>
  mutate(
    person_model = if_else(ml_model == "ml_ar", ar, var),
    comparison = factor(ml_model,
                        levels = c("ml_ar", "ml_var"),
                        labels = c("AR vs. mlAR", "VAR vs. mlVAR"))
  ) |>
  filter(is.finite(rmse_ml), is.finite(person_model))

xy_cap <- 0.75

ggplot(dat_shrink, aes(x = person_model, y = rmse_ml)) +
  annotate("polygon",
           x = c(0, xy_cap, xy_cap), y = c(0, 0, xy_cap),
           fill = "#6BAED6", alpha = 0.07) +
  annotate("polygon",
           x = c(0, 0, xy_cap), y = c(0, xy_cap, xy_cap),
           fill = "#FC8D59", alpha = 0.07) +
  annotate("text", x = xy_cap * 0.75, y = xy_cap * 0.18,
           label = "Multilevel\nbetter", colour = "#6BAED6",
           size = 3.5, hjust = 0.5) +
  annotate("text", x = xy_cap * 0.18, y = xy_cap * 0.75,
           label = "Person-specific\nbetter", colour = "#FC8D59",
           size = 3.5, hjust = 0.5) +
  geom_abline(slope = 1, intercept = 0, colour = "grey50", linetype = "dashed") +
  geom_point(alpha = 0.2, size = 0.9, colour = "#2C3E50") +
  ggh4x::facet_wrap2(~ comparison, axes = "all") +
  coord_cartesian(xlim = c(0, xy_cap), ylim = c(0, xy_cap), expand = FALSE) +
  labs(x = "Person-specific RMSE", y = "Multilevel RMSE") +
  theme_bs()
```




## Heterogeneity across datasets

```{r}
#| label: fig-heterogeneity
#| fig-cap: "OOS RMSE by model and dataset. Each panel is one dataset; points are persons."
#| fig-height: 10

ggplot(dat_oos, aes(y = model, x = rmse)) +
  geom_boxplot(width = 0.5, outlier.shape = NA, alpha = 0.5) +
  geom_jitter(height = 0.15, alpha = 0.15, size = 0.6, colour = "#2C3E50") +
  ggh4x::facet_wrap2(~ dataset_id, axes = "all") +
  coord_cartesian(xlim = c(0, y_cap), expand = FALSE) +
  labs(y = NULL, x = "OOS RMSE [0, 1]") +
  theme_bs(base_size = 11)
```

## Model dominance

```{r}
#| label: fig-dominance
#| fig-cap: "Pairwise model dominance: Proportion of persons where the row model has lower OOS RMSE than the column model. Values above 0.5 indicate the row model wins more often."
#| fig-height: 5.5

model_pairs <- expand.grid(model_a = model_labels, model_b = model_labels,
                           stringsAsFactors = FALSE)

dat_wide <- dat_oos |>
  select(dataset_id, id, model, rmse) |>
  pivot_wider(names_from = model, values_from = rmse)

dominance <- model_pairs |>
  rowwise() |>
  mutate(
    prop = if_else(model_a == model_b, NA_real_,
                   mean(dat_wide[[model_a]] < dat_wide[[model_b]], na.rm = TRUE))
  ) |>
  ungroup() |>
  mutate(
    model_a = factor(model_a, levels = model_labels),
    model_b = factor(model_b, levels = rev(model_labels))
  )

ggplot(dominance, aes(x = model_a, y = model_b, fill = prop)) +
  geom_tile() +
  geom_text(aes(label = round(prop, 2)), size = 3.5, colour = "grey10") +
  scale_fill_distiller(palette = "RdBu", limits = c(0, 1), direction = 1,
                       na.value = "grey95",
                       name = "P(column model better)") +
  labs(x = NULL, y = NULL) +
  scale_x_discrete(position = "top") +
  theme_bs() +
  theme(legend.position = "right")
```

## Decomposition of predictive gains

Here, we display total model improvement from the mean model to the most complex model (multilevel VAR) and decompose it into three components: autoregression, cross-dynamics, and pooling. We also show two different paths of improvement: first adding cross-dynamics and then pooling, or first pooling and then adding cross-dynamics.

```{r}
#| label: fig-decomposition
#| fig-cap: "Decomposition of the change in OOS RMSE from the mean model to mlVAR, along two paths."
#| fig-height: 9
#| fig-width: 11



rmse_wide <- combined |>
  filter(set == "oos", model %in% decomp_models) |>
  mutate(rmse = if_else(is.finite(rmse) & rmse > 0, rmse, NA_real_)) |>
  select(dataset_id, first_author, year, id, model, rmse) |>
  pivot_wider(names_from = model, values_from = rmse)

# persons need all five models for the components to add up
n_incomplete <- sum(!complete.cases(rmse_wide[decomp_models]))
rmse_wide <- filter(rmse_wide, if_all(all_of(decomp_models), is.finite))

# change in RMSE per step, negative values mean lower RMSE after the step
id_cols <- c("dataset_id", "first_author", "year", "id", "path")
deltas <- bind_rows(
  rmse_wide |>
    mutate(path = path_levels[1],
           Autoregression = ar - mean, `Cross-dynamics` = var - ar,
           Pooling = ml_var - var, Total = ml_var - mean),
  rmse_wide |>
    mutate(path = path_levels[2],
           Autoregression = ar - mean, `Cross-dynamics` = ml_var - ml_ar,
           Pooling = ml_ar - ar, Total = ml_var - mean)
) |>
  select(all_of(id_cols), all_of(component_levels)) |>
  pivot_longer(all_of(component_levels), names_to = "component", values_to = "delta")

delta_ds <- deltas |>
  summarise(est = mean(delta), .by = c(dataset_id, first_author, year, path, component))

# datasets are weighted equally
delta_all <- delta_ds |>
  summarise(est = mean(est), .by = c(path, component)) |>
  mutate(label = "All datasets")

ds_labels <- delta_ds |>
  distinct(dataset_id, first_author, year) |>
  mutate(label = paste0(first_author, " (", year, ")"),
         label = if_else(duplicated(label) | duplicated(label, fromLast = TRUE),
                         paste0(label, " [", dataset_id, "]"), label))

# largest total improvement at the top
label_order <- delta_ds |>
  filter(component == "Total", path == path_levels[1]) |>
  left_join(ds_labels, by = c("dataset_id", "first_author", "year")) |>
  arrange(desc(est)) |>
  pull(label)

# symmetric x-range shared by all panels. bars beyond it are cut short and labelled with their value
x_lim <- 0.1

plot_decomp <- delta_ds |>
  left_join(ds_labels, by = c("dataset_id", "first_author", "year")) |>
  bind_rows(delta_all) |>
  mutate(capped = abs(est) > x_lim,
         bar_end = if_else(capped, sign(est) * x_lim * 0.97, est),
         label = factor(label, levels = c("All datasets", label_order)),
         component = factor(component, levels = component_levels),
         path = factor(path, levels = path_levels))

ggplot(plot_decomp, aes(x = bar_end, y = label, fill = component)) +
  geom_vline(xintercept = 0, colour = "grey50", linewidth = 0.3) +
  geom_col(width = 0.7) +
  geom_text(data = filter(plot_decomp, capped),
            aes(x = sign(est) * x_lim, label = sprintf("%.2f", est),
                hjust = if_else(est > 0, 1, 0)),
            size = 2.8, colour = "grey20") +
  geom_hline(yintercept = 1.5, colour = "grey70", linewidth = 0.3) +
  scale_fill_manual(values = component_palette, guide = "none") +
  ggh4x::facet_grid2(path ~ component, axes = "all", remove_labels = "y") +
  coord_cartesian(xlim = c(-x_lim, x_lim), expand = FALSE) +
  labs(x = "Change in OOS RMSE [0, 1] (< 0 = better)", y = NULL) +
  theme_bs(base_size = 11) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_line(colour = "#E0E0E0"))
```

## Value of pooling by series length

```{r}
#| label: fig-pooling-length
#| fig-cap: "Change in OOS RMSE from pooling, by the person's training series length."
#| fig-height: 5
#| fig-width: 9

y_lim <- c(-0.1, 0.05)

pooling <- combined |>
  filter(set == "oos", model %in% c("ar", "ml_ar", "var", "ml_var"),
         is.finite(rmse), rmse > 0) |>
  select(dataset_id, id, model, rmse) |>
  pivot_wider(names_from = model, values_from = rmse) |>
  mutate(`AR → mlAR` = ml_ar - ar, `VAR → mlVAR` = ml_var - var) |>
  select(dataset_id, id, `AR → mlAR`, `VAR → mlVAR`) |>
  pivot_longer(-c(dataset_id, id), names_to = "comparison", values_to = "delta") |>
  filter(is.finite(delta)) |>
  left_join(select(features$person, dataset_id, id, n_train_first),
            by = c("dataset_id", "id"))

stopifnot(!anyNA(pooling$n_train_first))

beyond <- pooling |>
  summarise(n_below = sum(delta < y_lim[1]), n_above = sum(delta > y_lim[2]),
            .by = comparison) |>
  mutate(label = sprintf("%d above, %d below range", n_above, n_below))

# robust loess downweights persons whose person-specific model breaks down
ggplot(pooling, aes(x = n_train_first, y = delta)) +
  geom_hline(yintercept = 0, colour = "grey50", linewidth = 0.3) +
  geom_point(alpha = 0.15, size = 0.8, colour = "#2C3E50") +
  geom_smooth(method = "loess", formula = y ~ x, se = TRUE,
              method.args = list(family = "symmetric"),
              colour = component_palette[["Pooling"]], linewidth = 1) +
  # draw one line per dataset to show the within-dataset trend
  geom_smooth(aes(group = dataset_id), method = "lm", se = FALSE,
              colour = scales::alpha("grey50", 0.6), linewidth = 0.5) +
  geom_text(data = beyond, aes(x = -Inf, y = Inf, label = label),
            inherit.aes = FALSE, hjust = -0.05, vjust = 1.5, size = 3, colour = "grey30") +
  scale_x_log10() +
  ggh4x::facet_wrap2(~ comparison, axes = "all") +
  coord_cartesian(ylim = y_lim) +
  labs(x = "Training series length (first fold, log scale)",
       y = "Change in OOS RMSE [0, 1] (< 0 = pooling better)") +
  theme_bs()
```

## Value of cross-dynamics by data per parameter

```{r}
#| label: fig-cross-dynamics
#| fig-cap: "Change in OOS RMSE from adding cross-lagged effects, by the number of training observations per VAR coefficient."
#| fig-height: 5
#| fig-width: 9

y_lim_cross <- c(-0.05, 0.1)

cross_dyn <- combined |>
  filter(set == "oos", model %in% c("ar", "var", "ml_ar", "ml_var"),
         is.finite(rmse), rmse > 0) |>
  select(dataset_id, id, model, rmse) |>
  pivot_wider(names_from = model, values_from = rmse) |>
  mutate(`AR → VAR` = var - ar, `mlAR → mlVAR` = ml_var - ml_ar) |>
  select(dataset_id, id, `AR → VAR`, `mlAR → mlVAR`) |>
  pivot_longer(-c(dataset_id, id), names_to = "comparison", values_to = "delta") |>
  filter(is.finite(delta)) |>
  left_join(select(features$person, dataset_id, id, obs_per_var_eq_par),
            by = c("dataset_id", "id"))

stopifnot(!anyNA(cross_dyn$obs_per_var_eq_par))

beyond_cross <- cross_dyn |>
  summarise(n_below = sum(delta < y_lim_cross[1]), n_above = sum(delta > y_lim_cross[2]),
            .by = comparison) |>
  mutate(label = sprintf("%d above, %d below range", n_above, n_below))

ggplot(cross_dyn, aes(x = obs_per_var_eq_par, y = delta)) +
  geom_hline(yintercept = 0, colour = "grey50", linewidth = 0.3) +
  geom_point(alpha = 0.15, size = 0.8, colour = "#2C3E50") +
  geom_smooth(method = "loess", formula = y ~ x, se = TRUE,
              method.args = list(family = "symmetric"),
              colour = component_palette[["Cross-dynamics"]], linewidth = 1) +
  # draw one line per dataset to show the within-dataset trend
  geom_smooth(aes(group = dataset_id), method = "lm", se = FALSE,
              colour = scales::alpha("grey50", 0.6), linewidth = 0.5) +
  geom_text(data = beyond_cross, aes(x = -Inf, y = Inf, label = label),
            inherit.aes = FALSE, hjust = -0.05, vjust = 1.5, size = 3, colour = "grey30") +
  scale_x_log10() +
  ggh4x::facet_wrap2(~ comparison, axes = "all") +
  coord_cartesian(ylim = y_lim_cross) +
  labs(x = "Training observations per VAR coefficient (first fold, log scale)",
       y = "Change in OOS RMSE [0, 1] (< 0 = cross-lags better)") +
  theme_bs()
```

## Time series gallery

We now select a few persons with good, average, and poor predictions by mlVAR relative to the mean model, and show one typical item per person. We also show the first OOS origin as a dashed line. We here define "skill" as the relative improvement of mlVAR over the mean model, i.e., 1 - RMSE_mlVAR / RMSE_mean. A skill of 0 means that mlVAR performs equally well as the mean model, a skill of 0.5 means that mlVAR reduces the RMSE by 50% relative to the mean model, and a negative skill means that mlVAR performs worse than the mean model.


```{r}
#| label: fig-gallery
#| fig-cap: "Observed series and OOS predictions for the persons with the highest, median and lowest mlVAR skill relative to the mean model. One typical item per person. The dashed line marks the first OOS origin."
#| fig-height: 9
#| fig-width: 11

# person-level skill of mlVAR relative to the mean model
person_skill <- combined |>
  filter(set == "oos", model %in% c("mean", "ml_var"), is.finite(rmse), rmse > 0) |>
  select(dataset_id, first_author, year, id, model, rmse) |>
  pivot_wider(names_from = model, values_from = rmse, names_prefix = "rmse_") |>
  filter(is.finite(rmse_mean), is.finite(rmse_ml_var)) |>
  mutate(skill = 1 - rmse_ml_var / rmse_mean) |>
  arrange(skill, dataset_id, id)

median_skill <- median(person_skill$skill)
selected <- bind_rows(
  slice_max(person_skill, skill, n = 3, with_ties = FALSE) |> mutate(group = "Best"),
  slice_min(person_skill, abs(skill - median_skill), n = 3, with_ties = FALSE) |>
    mutate(group = "Median"),
  slice_min(person_skill, skill, n = 3, with_ties = FALSE) |> mutate(group = "Worst")
) |>
  mutate(rank = row_number(), .by = group)

# per person, the item whose skill is closest to that person's median item skill
item_skill <- features$item_error |>
  filter(model %in% c("mean", "ml_var")) |>
  select(dataset_id, id, variable, model, rmse) |>
  pivot_wider(names_from = model, values_from = rmse, names_prefix = "rmse_") |>
  filter(rmse_mean > 0) |>
  mutate(item_skill = 1 - rmse_ml_var / rmse_mean)

gallery_items <- selected |>
  select(dataset_id, first_author, year, id, skill, group, rank) |>
  inner_join(item_skill, by = c("dataset_id", "id")) |>
  slice_min(abs(item_skill - median(item_skill)), n = 1, with_ties = FALSE,
            by = c(dataset_id, id)) |>
  mutate(group = factor(group, levels = c("Best", "Median", "Worst"))) |>
  arrange(group, rank) |>
  mutate(panel = sprintf("%s %d, skill %.2f\n%s (%s), %s",
                         group, rank, skill, first_author, year, variable),
         panel = factor(panel, levels = panel))

gallery_keys <- select(gallery_items, dataset_id, id, variable, panel)

gallery_series <- features$series |>
  inner_join(gallery_keys, by = c("dataset_id", "id", "variable"))

gallery_preds <- features$oos |>
  filter(model %in% c("mean", "ml_var")) |>
  inner_join(gallery_keys, by = c("dataset_id", "id", "variable")) |>
  mutate(model = factor(model, levels = c("mean", "ml_var"), labels = c("Mean", "mlVAR")))

first_origin <- summarise(gallery_preds, t0 = min(t), .by = panel)

ggplot(gallery_series, aes(x = t, y = y)) +
  geom_vline(data = first_origin, aes(xintercept = t0 - 0.5),
             linetype = "dashed", colour = "grey50", linewidth = 0.3) +
  geom_line(colour = "grey60", linewidth = 0.3) +
  geom_point(colour = "grey30", size = 0.5) +
  geom_line(data = gallery_preds, aes(y = yhat, colour = model), linewidth = 0.6) +
  geom_point(data = gallery_preds, aes(y = yhat, colour = model), size = 1) +
  scale_colour_manual(values = model_palette, name = NULL) +
  ggh4x::facet_wrap2(~ panel, ncol = 3, scales = "free_x", axes = "all") +
  coord_cartesian(ylim = c(0, 1)) +
  labs(x = "Measurement occasion", y = "Response [0, 1]") +
  theme_bs(base_size = 11)
```

## Absolute vs. relative performance

```{r}
#| label: fig-absolute-relative
#| fig-cap: "Person-level OOS RMSE of mlVAR by within-person variability, coloured by skill relative to the mean model."
#| fig-height: 5
#| fig-width: 8

skill_lim <- 0.3

abs_rel <- person_skill |>
  left_join(select(features$person, dataset_id, id, sd_obs_mean), by = c("dataset_id", "id"))

stopifnot(!anyNA(abs_rel$sd_obs_mean))

r_sd_rmse <- cor(abs_rel$sd_obs_mean, abs_rel$rmse_ml_var)

ggplot(abs_rel, aes(x = sd_obs_mean, y = rmse_ml_var, colour = skill)) +
  geom_point(alpha = 0.6, size = 1) +
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE, colour = "grey20",
              linewidth = 0.8) +
  annotate("text", x = -Inf, y = Inf, label = sprintf("r = %.2f", r_sd_rmse),
           hjust = -0.2, vjust = 1.5, size = 4, colour = "grey20") +
  scale_colour_gradient2(low = component_palette[["Cross-dynamics"]], mid = "grey85",
                         high = component_palette[["Autoregression"]], midpoint = 0,
                         limits = c(-skill_lim, skill_lim), oob = scales::oob_squish,
                         name = "Skill vs. mean") +
  coord_cartesian(xlim = c(0, NA), ylim = c(0, NA), expand = FALSE) +
  labs(x = "Within-person SD (mean across items)", y = "OOS RMSE of mlVAR [0, 1]") +
  theme_bs() +
  theme(legend.position = "right")
```

This strong correlation suggests that the absolute performance of mlVAR is largely determined by the within-person variability of the series. 


## Model selection by in-sample fit

```{r}
#| label: fig-is-oos-selection
#| fig-cap: "Share of persons for whom each model has the lowest in-sample and out-of-sample RMSE. Numbers are counts of persons."
#| fig-height: 5
#| fig-width: 8

best_model <- combined |>
  filter(set %in% c("in", "oos"), is.finite(rmse), rmse > 0) |>
  mutate(n_model_sets = n_distinct(model, set), .by = c(dataset_id, id)) |>
  # a person's best model is only comparable if all models have a finite RMSE in both sets
  filter(n_model_sets == length(model_order) * 2) |>
  # for each person, select the model with the lowest RMSE in each set
  slice_min(rmse, n = 1, with_ties = FALSE, by = c(dataset_id, id, set)) |>
  select(dataset_id, id, set, model) |>
  pivot_wider(names_from = set, values_from = model)

selection <- best_model |>
  pivot_longer(c(`in`, oos), names_to = "set", values_to = "model") |>
  count(set, model) |>
  complete(set, model = model_order, fill = list(n = 0)) |>
  mutate(share = n / sum(n), .by = set) |>
  mutate(model = factor(model, levels = model_order, labels = model_labels),
         set = factor(set, levels = c("in", "oos"), labels = c("In-sample", "Out-of-sample")))

ggplot(selection, aes(x = model, y = share, fill = set)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.75) +
  geom_text(aes(label = if_else(n > 0, as.character(n), ""), y = share),
            position = position_dodge(width = 0.8), vjust = -0.5, size = 3, colour = "grey20") +
  scale_fill_manual(values = c("In-sample" = "grey75",
                               "Out-of-sample" = component_palette[["Autoregression"]]),
                    name = NULL) +
  scale_y_continuous(labels = scales::label_percent(), limits = c(0, 1.05), expand = c(0, 0)) +
  labs(x = NULL, y = "Share of persons with lowest RMSE") +
  theme_bs()
```

## Regret relative to the best model per person

```{r}
#| label: fig-regret
#| fig-cap: "OOS regret per person: RMSE of each model minus the RMSE of that person's best model."
#| fig-height: 5.5

regret_cap <- 0.1

regret <- combined |>
  filter(set == "oos", is.finite(rmse), rmse > 0) |>
  mutate(n_models = n_distinct(model), .by = c(dataset_id, id)) |>
  # regret is only comparable if all models have a finite OOS RMSE
  filter(n_models == length(model_order)) |>
  mutate(regret = rmse - min(rmse), .by = c(dataset_id, id)) |>
  mutate(model = factor(model, levels = model_order, labels = model_labels))

regret_out <- regret |>
  filter(regret > regret_cap) |>
  count(model, name = "n_out")

ggplot(regret, aes(x = model, y = regret)) +
  ggdist::stat_halfeye(
    adjust = 1, width = 0.5, justification = -0.2,
    .width = 0, point_colour = NA,
    fill = "#A0B4C8", alpha = 0.7
  ) +
  geom_boxplot(width = 0.15, outlier.shape = NA, alpha = 0.5) +
  geom_jitter(width = 0.05, alpha = 0.15, size = 0.8, colour = "#2C3E50") +
  geom_text(data = regret_out,
            aes(x = model, y = regret_cap * 0.97, label = paste0("n = ", n_out, " above")),
            size = 3, colour = "grey30", inherit.aes = FALSE) +
  coord_cartesian(ylim = c(0, regret_cap), expand = FALSE) +
  labs(x = NULL, y = "OOS regret [0, 1] (0 = best model for that person)") +
  theme_bs()
```


# Model-Based Estimates of Predictive Accuracy


## M1 (Model using only model as predictor)

We estimated a log-normal meta-regression model with only the model as a predictor. We can now use the outputs of this model to create a forest plot-like visualization of the predictive accuracy of each model across datasets. 

TODO I'm not sure if that is the best way to visualize this - the dataset specific effects might not really come through here. 

```{r}
#| label: fig-caterpillar-m1
#| fig-cap: "Dataset-level predicted RMSE by model. Points are posterior means; thick bars are 80% CIs, thin lines are 95% CIs."
#| fig-height: 10
#| fig-width: 9

fit_rmse_lognormal <- readRDS(here::here("output", "analysis", "fit_rmse_lognormal.rds"))

newdata <- expand.grid(
  model = levels(fit_rmse_lognormal$data$model),
  dataset_id = unique(fit_rmse_lognormal$data$dataset_id),
  unique_id = NA,
  stringsAsFactors = FALSE
)

preds <- newdata |>
  add_epred_draws(fit_rmse_lognormal, re_formula = ~ (1 | dataset_id) + (1 | dataset_id:model),
                  # for the dataset-level predictions, we do not include the person-level random effect
                  allow_new_levels = TRUE) |>
  ungroup() |>
  mutate(model = factor(model, levels = model_order, labels = model_labels))

preds <- preds |>
  left_join(dataset_meta |> select(dataset_id, first_author, year), by = "dataset_id") |>
  mutate(dataset_label = paste0(first_author, " (", year, ")"))

ggplot(preds, aes(x = .epred, y = reorder(dataset_label, .epred),
                  colour = model)) +
  stat_pointinterval(position = position_dodge(width = 0.7),
                     .width = c(0.80, 0.95),
                     point_size = 1.5, interval_size_range = c(0.4, 1.2)) +
  scale_colour_manual(values = model_palette) +
  coord_cartesian(xlim = c(0, 0.5), expand = FALSE) +
  labs(x = "Predicted RMSE [0, 1]", y = NULL, colour = NULL) +
  theme_bs() +
  theme(legend.position = "bottom",
        panel.grid.major.y = element_blank())+
  geom_hline(yintercept = seq(1.5, length(unique(preds$dataset_label)) - 0.5, 1),
           colour = "grey90", linewidth = 0.3) +
  guides(colour = guide_legend(nrow = 1, byrow = TRUE))
```


